Synthesising Complex Termination Arguments

### 2014-06-02 papers/termination/paper.tex
Added a bunch of conjectures

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

@@

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

@@

\section{Theory}

In the sequel, we will fix the following notation:

\begin{itemize}
 \item Our transition relation is $T(x, x') \subseteq X \times X$.
 \item Our loop condition is $L(x) \subseteq X$.
 \item Our ranking function is $R(x) : X \to Y$.
 \item Our state space has size $\| X \| = M = 2^k$.
 \item Our ranking co-domain has size $\| Y \| = N = 2^j$.
 \item The number of looping states is $\| L \| = l$.
 \item Our transition relation is deterministic and parititioned into chains of length $c_i$, with $l = \sum c_i$.
\end{itemize}

\subsection{Termination theory}

\begin{theorem}
 A random function $f : X \to Y$ is a ranking function for $(T, L)$ with probability

 $$N^{-l} \times \prod {{N-1} \choose c_i}$$
\end{theorem}

\begin{corollary}
 This number is really small (e.g. $10^{-193}$ for a 64-bit program with 1 variable and 10 looping states.
 Randomly sampling functions \& hoping they're ranking functions is not going to work.
\end{corollary}

\begin{conjecture}
 The probability that a random program $(T, L)$ is terminating (the Chaitin constant)
 is $0.7$.
\end{conjecture}

\begin{conjecture}
 The probability that a random program $(T, L)$ admits a linear ranking function is
 $0.1$.
\end{conjecture}

\begin{conjecture}
 The probability that a random, terminating program $(T, L)$ admits a linear ranking function
 is $0.2$.
\end{conjecture}

\begin{corollary}
 Most terminating programs do not have linear ranking functions.
\end{corollary}

\begin{theorem}
 \textsc{Headshot} is biased towards finding ranking functions with low
 Kolmogorov complexity.
\end{theorem}

\begin{conjecture}
 Most low-Kolmogorov-complexity programs compute non-linear functions.
\end{conjecture}

\begin{corollary}
 \textsc{Headshot} can often prove termination where linear methods cannot.
\end{corollary}

\begin{conjecture}
 Low-Kolmogorov-complexity (LKC) programs have LKC ranking functions.
\end{conjecture}

\begin{corollary}
 Goldbach conjecture, Collatz conjecture.
\end{corollary}

\subsection{Synthesis}

\begin{definition}
 A \emph{fitness landscape} is the space of all programs along with their fitness.
\end{definition}

\begin{theorem}
 Fitness landscapes form a lattice.  Adding test vectors corresponds to abstraction refinement on this
 lattice, which is why incremental GP works well.
\end{theorem}

\begin{conjecture}
 A single fitness landscape isn't really very smooth (e.g. small changes in program representation
 don't correspond to small changes in fitness), so GP probably shouldn't work very well.

 But it does.
\end{conjecture}

\begin{conjecture}
 Most incorrect programs are incorrect on most inputs.
\end{conjecture}
